---
title: 'Final Project'
author: "Yigit Tahmisoglu"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Github repository
https://github.com/tahmisoglu-yigit/R_Project.git

## Dataset

```{r}
#tinytex::install_tinytex()
data <- read.csv("C:/Users/YIGIT/R/TRIR71.csv")


```

## Description of dataset

The data set is the one I plan to use for my masters thesis.*(They say not to share the dataset with someone else, but I don't think there will be an issue to share with you.)*
It is a cross-sectional survey data from the Demographic and Health Surveys for Turkey from the year 2018, which is a representative household  survey providing detailed information on birth recodes and individual records for women of each household for developing countries. 

The potential questions I am interested in exploring are:

-   Effect of childbirth on female labor force participation

-   Effect of living conditions on early childhood development

-   And additional ones I might discover along the way


