---
title: 'Final Project'
author: "Yigit Tahmisoglu"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Github repository
https://github.com/tahmisoglu-yigit/R_Project.git

## Trying to open this friggin dataset

```{r}

library(rio)

# Data is in dropbox, let's retrieve it.
dropbox <- "https://www.dropbox.com/s/zemcx6rqxsxu9ht/TRIR71.csv?dl=1"
df <- import(dropbox, format = "csv")
# Voila

View(df)
ls(df)
sapply(df, mean, na.rm=TRUE)
summary(df)

```

## Description of dataset

The data set is the one I plan to use for my masters thesis.
It is a cross-sectional survey data from the Demographic and Health Surveys for Turkey from the year 2018, which is a representative household  survey providing detailed information on birth recodes and individual records for women of each household for developing countries. 

The potential questions I am interested in exploring are:

-   Effect of childbirth on female labor force participation



